Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeast Taiwan

### 2019-10-29 analysis/paper/paper.Rmd
added code chucks for reading data and preparing data for EFA

```{r}
# read in data
# EFA
```

@@

>>>>

@@

```{r read-data}
library(here)
lf <- list.files(here("analysis", "data", "raw_data"), full.names = TRUE)

library(xml2)
library(stringr)
remove_scan_and_fill <- function(svg_file) {
  # remove the scan image from the SVG
  t1 <- read_xml(svg_file)
  t2 <- xml_children(t1)
  scan_layer <- t2[[4]]
  xml_remove(scan_layer)
  # now we have a SVG without the scan layer

  # change from outline to fill
  stroke_and_fill_node <- xml_children(t1)[4]
  stroke_and_fill_node_inner <- xml_children(stroke_and_fill_node)
  xml_attr(stroke_and_fill_node_inner, "style") <- 
    "fill:#000000;stroke:none;stroke-width:5;stroke-linecap:butt;stroke-linejoin:miter;stroke-miterlimit:4;stroke-dasharray:none;stroke-opacity:1"

  write_xml(t1,
            str_glue("analysis/data/raw_data/fill_{basename(svg_file)}"))

  # fit image object to canvas in svg so it's all captured during the conversion to JPG
  system(str_glue('inkscape --verb=EditSelectAll --verb=AlignHorizontalCenter --verb=AlignVerticalCenter  --verb=FitCanvasToSelectionOrDrawing --verb=FileSave --verb=FileQuit "$(pwd)/analysis/data/raw_data/fill_{basename(svg_file)}"'))
}

# run the function on all SVGs
library(purrr)
walk(lf, remove_scan_and_fill)

# convert SVG to PNG
filled_svgs <- list.files("analysis/data/raw_data", pattern = "^fill_", full.names = TRUE)

# convert svgs to to jpg
library(magick)
library(rsvg)
map(filled_svgs,
  ~image_read_svg(.x) %>% 
  image_background("white") %>%  
  image_convert("jpg") %>% 
  image_write(str_glue("{.x}_output.jpg")))
```

```{r tidy-data}
# Once we have the jpgs, start from here
library(tidyverse)

# get list of jpgs only
jpgs <- list.files("analysis/data/raw_data", pattern = ".jpg", full.names = TRUE)

# join file names to phase
pottery_data <- readr::read_csv("Pottery_outlines/KWL_Ceramic_final2018_with_phases.csv")
pottery_data <- pottery_data %>% 
  mutate(filename = str_glue('fill_{filename}.svg_output.jpg'))

# some outlines that we have no phases for... 
sdiff <- setdiff(pottery_data$filename, basename(jpgs))

# the ones we do have both outlines and phases
isect <- intersect(pottery_data$filename, basename(jpgs))

jpgs <- jpgs[basename(jpgs) %in% isect]
pottery_data <- 
  pottery_data %>% 
  filter(filename %in% isect) %>% 
  distinct(filename, phase, .keep_all = TRUE)

phases_isect <- 
  pottery_data %>%  
  select(filename, phase) %>% 
  filter(filename %in% isect) %>% 
  mutate_all(as.factor) %>% 
  distinct(filename, phase) %>% 
  mutate(phase = as.factor(case_when(
    phase == 'post-e' ~ "Post-European",
    phase == 'pre-e' ~ "Pre-European",
    phase == 'ch-con' ~ "Post-European" # we combine Chinese and post-Euro
  )))

# how many pots in each phase?
phases_isect %>% 
  group_by(phase) %>% 
  tally(sort = T)

# start shape exploration
library(Momocs)
#  it freezes on the empty ones...
jpgs_imported <- import_jpg(jpgs, threshold = 0.5)

# Normalise the shapes, from 
# https://www.repository.cam.ac.uk/bitstream/handle/1810/266423/Supplementary%20Code%20S1.pdf?sequence=20&isAllowed=y
jpgs_imported_coo <- 
  jpgs_imported %>% 
   Out(., fac = phases_isect)  %>% 
  # Centring, scaling and sampling pseudo-landmarks
   coo_center %>% 
   coo_scale %>% 
   coo_sample(., 1000) %>% 
  # Procrustes superimposition
   fgProcrustes %>% 
  # Normalisation of the starting points
   coo_slidedirection

jpgs_imported_coo %>% 
  panel(., fac = phases_isect$phase)

pile(jpgs_imported_coo)

```
